Adversarial Variational Optimization of\\ Non-Differentiable Simulators

### 2018-01-30 icml/avo.tex
save

\begin{abstract}
    Complex computer simulators are increasingly used across fields of science as
    generative models tying parameters of an underlying theory to
    experimental observations. Inference in this setup is often
    difficult, as simulators rarely admit a tractable density or likelihood
    function. We introduce Adversarial Variational Optimization (AVO), a likelihood-free
    inference algorithm for fitting a non-differentiable generative model incorporating ideas from empirical Bayes and variational inference.
    We adapt the training procedure of generative
    adversarial networks by replacing the differentiable generative network with a
    domain-specific simulator. We solve the resulting non-differentiable
    minimax problem by minimizing variational upper bounds of the two adversarial objectives.
    Effectively, the procedure results in learning a
    proposal distribution over simulator parameters, such that the corresponding
    marginal distribution of the generated data matches the observations.
    We present results of the method with simulators producing both discrete and continuous data.
\end{abstract}

@@

We assume that the stochastic generative process that defines $p(\mathbf{x}|\bftheta)$ is
specified through a non-differentiable deterministic function $g(\cdot; \bftheta) : {\cal Z} \to
\mathbb{R}^d$. Operationally, %That is, we consider
\begin{equation}\label{eqn:p_theta}
    \mathbf{x} \sim p(\mathbf{x}|\bftheta) \equiv \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta)
\end{equation}
such that the density $p(\mathbf{x}|\bftheta)$ can be
written as
\begin{equation}\label{eqn:p_x_sim}
    p(\mathbf{x}|\bftheta) = \int_{\{\bfz:g(\bfz;\bftheta) = \bfx \}} p(\bfz|\bftheta) \mu(d\bfz),
\end{equation}
where $\mu$ is a probability measure.

@@

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful in many problems.
We use Adam~\cite{2014arXiv1412.6980K}, but note the opportunity to use instead the Natural Evolution Strategy
algorithm~\cite{2011arXiv1106.4487W} or variance reduction techniques
such as Rao-Blackwellization,
control variates, or quasi-Monte Carlo~\cite{ranganath2014black, tran2017variational}.

@@

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the parametrized marginal distribution of the generated data
\begin{equation}\label{eq:marginal_likelihood}
\qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \cite{2016arXiv161009033R}, though more complicated than a simple
reparametrization of normally distributed noise $\bfz$ through a differentiable
function. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ whose expectations are differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
\mathbf{x} \sim \qxpsi \equiv \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

In order to more effectively target
point estimates $\bftheta^*$, we augment Eqn.~\ref{eqn:vo-ug} with
an entropic regularization term $H(q(\bftheta|\bfpsi))$, that is
\begin{equation}
U_g = \mathbb{E}_{\bftheta \sim q(\bftheta|\bfpsi)} [ {\cal L}_g ] + \gamma H(q(\bftheta|\bfpsi))
\end{equation}
where $\gamma \in \mathbb{R}^+$ is a hyper-parameter controlling the trade-off
between the generator objective and the tightness of the proposal distribution.
For small values of $\gamma$,
proposal distributions with large entropy are not penalized, which results
in learning a smeared variation of the original simulator corresponding to the marginal or prior predictive model. On the other hand,
for large values of $\gamma$, the procedure is encouraged to fit a proposal
distribution with low entropy, which has the effect of concentrating its density
tightly around one or a few $\bftheta$ values.

@@

@@

The observed data is sampled from a Poisson with mean $\lambda^* = 7$.
Algorithm~\ref{alg:avo} is run for 300 epochs with mini-batches of size $M=64$
and the following configuration. For the critic $d$, we use a 3-layer MLP with 10
hidden nodes per layer and ReLU activations. At each epoch, Adam is run for
$n_\text{critic}=100$ iterations with a step size $\alpha=0.01$, decay rates
$\beta_1=\beta_2=0.5$ and its inner first and second moment vectors reset at
each outer epoch in order to avoid building momentum in staled directions.  For
estimating $\lambda^*$, we parametrize $\bftheta$ as $\log(\lambda)$ and use a univariate Gaussian proposal distribution
$q(\bftheta|\bfpsi)$ initialized with a mean at $\log(5)$ and unit variance. At
each epoch, parameters $\bfpsi$ are updated by taking one Adam step, with
$\alpha=0.01$ and $\beta_1=\beta_2=0.5$. The gradient penalty coefficient is set to
$\lambda=0.025$, and the entropy penalty is evaluated at both $\gamma=0$ and $\gamma=5$.

The top left plot in Figure~\ref{fig:poisson} illustrates the resulting proposal
distributions $q(\bftheta|\bfpsi)$ after AVO.  For
both $\gamma=0$ and $\gamma=5$, the proposal distributions correctly concentrate
their density around the true parameter value $\log(\lambda^*) = 1.94$. Under
the effect of the positive entropy penalty $H(q(\bftheta|\bfpsi))$,
the proposal distribution for $\gamma=5$ concentrates its mass more tightly,
yielding in this case a more precise inference.  The top right plot compares the
model distributions to the true distribution.  As theoretically expected from
adversarial training, we see that the resulting distributions closely match
the true distribution, with in this case visually slightly better results for the penalized
model.  The bottom plot of Figure~\ref{fig:poisson} shows empirical estimates
of $-U_d$ with respect to the epoch number. For both $\gamma=0$ and $\gamma=5$,
the curves quickly fall towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\bftheta)}
[d(\tilde{\mathbf{x}};\bfphi)] \approx \mathbb{E}_{\mathbf{x} \sim
p_r(\mathbf{x})} [d(\mathbf{x};\bfphi)]$ and that the critic cannot distinguish
between true and model data. This confirms that adversarial variational optimization
works despite the discreteness of the data and lack of access to the
density $p(\bfx | \bftheta)$ or its gradient.

\begin{figure}[h]
\centering
\includegraphics[width=0.48\textwidth]{figures/poisson.pdf}
\caption{Discrete Poisson model with unknown mean.
 ({\it Top left}) Proposal distributions $q(\bftheta|\bfpsi)$ after training. For both $\gamma=0$ and $\gamma=5$, the distributions correctly concentrate their density around
            the true value $\log(\lambda^*)$. Penalizing the entropy of the proposal distribution ($\gamma=5$) results in a tighter density.
 ({\it Top right}) Model distributions $\qxpsi$ after training. This plot shows that the resulting parametrizations of the simulator closely reproduce the true distribution.
 ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
 }\label{fig:poisson}
\end{figure}

@@

As a second toy example, we consider a generator producing
5-dimensional continuous data, as originally specified in Section 4.2. of
\cite{cranmer2015approximating}. More specifically, we consider the following
generative process:
\begin{itemize}
\item $\bfz = (z_0, z_1, z_2, z_3, z_4)$, such that \\
$z_0 \sim {\cal N}(\mu=\alpha, \sigma=1)$,
$z_1 \sim {\cal N}(\mu=\beta, \sigma=3)$,
$z_2 \sim {\text{Mixture}}[\frac{1}{2}\,{\cal N}(\mu=-2, \sigma=1), \\  \frac{1}{2}\,{\cal N}(\mu=2, \sigma=0.5)]$,
$z_3 \sim {\text{Exponential}(\lambda=3)}$, and
$z_4 \sim {\text{Exponential}(\lambda=0.5)}$;
\item $\mathbf{x} = R  \bfz$, where $R$ is a fixed
semi-positive definite $5 \times 5$ matrix defining a fixed projection
of $\bfz$ into the observed space.
\end{itemize}
Again, the AVO algorithm does not have access to the density or its gradient, only samples from the generative model.
We consider observed data generated at the nominal values $\bftheta^* = (\alpha^*=1,\beta^*=-1)$.
The simulator parameters are modeled with a factored  (mean field) Gaussian
proposal distribution $q(\bftheta|\bfpsi) = q(\alpha|\bfpsi) q(\beta|\bfpsi)$, where each component was
initialized with zero mean and unit variance.
Hyper-parameters are set to $M=64$, $n_\text{critic}=100$, $\lambda=0.025$, $\gamma=10$ and
Adam configured with $\alpha=0.01$, $\beta_1=0.9$ and $\beta_2=0.999$.
The architecture of the critic is the same as in the previous example.

Starting with a proposal distribution $q(\bftheta|\bfpsi)$ largely spread over
the parameter space, as illustrated in the left plot of Figure~\ref{fig:multi},
AVO quickly converges towards a variational distribution whose density
concentrates around the nominal values $\bftheta^*$, as shown in the right plot of Figure~\ref{fig:multi}.
Overall, this example further illustrates and confirms the ability of adversarial
variational optimization for inference with multiple parameters and multidimensional
data, where reliable approximations of $p(\mathbf{x}|\bftheta)$ in a traditional
MLE setting would otherwise be difficult to construct.

@@

@@

This work sits at the intersection of several lines of research related to
likelihood-free inference, approximate Bayesian computation (ABC),
implicit generative networks, and variational inference.
Viewed from the literature around implicit generative models based on neural networks,
the proposed method can be considered as a direct adaptation of
generative adversarial networks~\cite{goodfellow2014generative,2017arXiv170104862A, 2017arXiv170107875A} to
non-differentiable simulators using variational optimization~\cite{2012arXiv1212.4507S,staines2013optimization}.

@@

Likewise, AVO closely relates to recent extensions of GANs, such as
Adversarial Learned Inference (ALI)~\cite{dumoulin2016adversarially},
Bidirectional GANs (BiGANs)~\cite{donahue2016adversarial},
$\alpha$-GAN~\cite{rosca2017variational}, Adversarial Variational Bayes
(AVB)~\cite{DBLP:journals/corr/MeschederNG17}, and the \texttt{PC-Adv}
algorithm of  \cite{2017arXiv170208235H}, which
add an inference network to the generative model.  Each of these assume a
tractable density $p(\bfx|\bftheta)$ that is differentiable with respect to
$\bftheta$, which is  not satisfied in the likelihood-free setting. Our lifting
of the non-differentiable simulator $p(\bfx|\bftheta)$ to the variational
program $q(\bfx | \bfpsi)$ provides the ability to differentiate expectations
with respect to $\bfpsi$ as in Eqn~\ref{eqn:approx-grad}; however, the density
$q(\bfx | \bfpsi)$ is still intractable. Moreover, we do not attempt to define a
recognition model  $q(\bfz, \bftheta|\bfpsi)$ as the latent space $\mathcal{Z}$
of many real-world simulators is complicated and not amenable to neural a
recognition model.

@@

Preliminary results on toy problems with discrete and continuous data validate
the proposed method. While the obtained results are encouraging, the complete
validation of the method remains to be carried out in real conditions on a full
fledged scientific simulator -- which we plan to achieve for a next version of
this work. In terms of method, several components need further investigation.
First, we need to better study the interplay between the entropy penalty and the
adversarial objectives. Second, we should better understand the dynamics of the
optimization procedure, in particular when combined with momentum-based
optimizers like Adam. Third, we need to consider whether less noisy estimates of
the gradients $\nabla_\bfpsi U_g$ can be computed.

@@

As a more realistic example, we now consider a (simplified) simulator from
particle physics for electron--positron collisions resulting in muon--antimuon
pairs ($e^+e^- \rightarrow \mu^+\mu^-$). The simulator approximates the
distribution of observed measurements $\mathbf{x} = \cos(A) \in [-1,1]$, where $A$ is the
polar angle of the outgoing muon with respect  to the originally incoming
electron. Neglecting measurement uncertainty induced from the particle detectors,
this random variable is approximately distributed as
\begin{equation}\label{eq:simple_weinberg}
p(\mathbf{x}|E_\text{beam}, G_f) = \frac{1}{Z} \left[ (1 + \mathbf{x}^2) + c(E_\text{beam}, G_f) \mathbf{x} \right]
\end{equation}
where $Z$ is a known normalization constant and $c$ is an asymmetry coefficient
function. Due to the linear term in the expression, the density $p(\mathbf{x} |
E_\text{beam}, G_f)$ exhibits a so-called {\it forward-backward} asymmetry.  Its
size depends on the values of the parameters $E_\text{beam}$ (the beam energy)
and $G_f$ (the Fermi constant) through the coefficient function $c$.

>>>>

\begin{abstract}
    Complex computer simulators are increasingly used across fields
of science as generative models tying parameters of an underlying theory to
experimental observations. Inference in this setup is often difficult, as
simulators rarely admit a tractable density or likelihood function. We introduce
Adversarial Variational Optimization (AVO), a likelihood-free inference
algorithm for fitting a non-differentiable generative model incorporating ideas
from generative adversarial networks (GANs) and variational inference. We adapt the
training procedure of (Wasserstein) GANs by replacing the
differentiable generative network with a domain-specific simulator. We solve the
resulting non-differentiable minimax problem by minimizing variational upper
bounds of the two adversarial objectives. Effectively, the procedure results in
learning a proposal distribution over simulator parameters, such that the
corresponding marginal distribution of the generated data matches the
observations. We present results of the method with simulators producing both
discrete and continuous data.
\end{abstract}

@@

We assume that the stochastic generative process that defines $p(\mathbf{x}|\bftheta)$ is
specified through a non-differentiable deterministic function $g(\cdot; \bftheta) : {\cal Z} \to
\mathbb{R}^d$. Operationally, %That is, we consider
\begin{equation}\label{eqn:p_theta}
    \mathbf{x} \sim p(\mathbf{x}|\bftheta) \triangleq \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta)
\end{equation}
such that the density $p(\mathbf{x}|\bftheta)$ can be
written as
\begin{equation}\label{eqn:p_x_sim}
    p(\mathbf{x}|\bftheta) = \int_{\{\bfz:g(\bfz;\bftheta) = \bfx \}} p(\bfz|\bftheta) \mu(d\bfz),
\end{equation}
where $\mu$ is a probability measure.

@@

Algorithm~\ref{alg:avo} represents the simplest version of AVO; however, the
variance of the noisy estimator of the gradients may be too large to be useful in many problems.
We use Adam~\cite{2014arXiv1412.6980K}, but note the opportunity to use instead the Natural Evolution Strategy
algorithm~\cite{2011arXiv1106.4487W} or variance reduction techniques as
control variates~\cite{2017arXiv171100123G} or finite differences~\cite{buesingstochastic}.

@@

The variational objectives \ref{eqn:vo-ud}-\ref{eqn:vo-ug}
have the effect of replacing the modeled data distribution of Eqn.~\ref{eqn:p_theta} with
the parametrized marginal distribution of the generated data
\begin{equation}\label{eq:marginal_likelihood}
\qxpsi = \int  p(\mathbf{x}|\bftheta) q(\bftheta|\bfpsi) d\bftheta.
\end{equation}
We can think of $q(\bfx|\bfpsi)$ as a \textit{variational program} as described
in \cite{2016arXiv161009033R}, though more complicated than a simple
reparametrization of normally distributed noise $\bfz$ through a differentiable
function. In our case, the variational program is a
marginalized, non-differentiable  simulator.  Its density is  intractable;
nevertheless, it can generate samples for $\bfx$ whose expectations are differentiable with
respect to $\bfpsi$.
Operationally, we sample from this marginal model via
\begin{equation}\label{eqn:p_psi}
\mathbf{x} \sim \qxpsi \triangleq \bftheta \sim q(\bftheta|\bfpsi), \bfz \sim p(\bfz|\bftheta), \mathbf{x} = g(\bfz; \bftheta).
\end{equation}

@@

In order to more effectively target
point estimates $\bftheta^*$, we augment Eqn.~\ref{eqn:vo-ug} with
an entropic regularization term $H(q(\bftheta|\bfpsi))$, that is
\begin{equation}
U_g = \mathbb{E}_{\bftheta \sim q(\bftheta|\bfpsi)} [ {\cal L}_g ] + \gamma H(q(\bftheta|\bfpsi))
\end{equation}
where $\gamma \in \mathbb{R}^+$ is a hyper-parameter controlling the trade-off
between the generator objective and the tightness of the proposal distribution.
For small values of $\gamma$,
proposal distributions with large entropy are not penalized, which results
in learning a smeared variation of the original simulator. On the other hand,
for large values of $\gamma$, the procedure is encouraged to fit a proposal
distribution with low entropy, which has the effect of concentrating its density
tightly around one or a few $\bftheta$ values.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.48\textwidth]{figures/poisson.pdf}
\caption{Discrete Poisson model with unknown mean.
 ({\it Top left}) Proposal distributions $q(\bftheta|\bfpsi)$ after training. For both $\gamma=0$ and $\gamma=1$, the distributions correctly concentrate their density around
            the true value $\log(\lambda^*)$. Entropic regularization ($\gamma=1$) results in a tighter density.
 ({\it Top right}) Model distributions $\qxpsi$ after training. This plot shows that the resulting parametrizations of the simulator closely reproduce the true distribution.
 ({\it Bottom}) Empirical estimates of the variational upper bound $U_d$ as optimization progresses.
 }\label{fig:poisson}
\end{figure}

@@

The observed data is sampled from a Poisson with mean $\lambda^* = 7$.
Algorithm~\ref{alg:avo} is run for 500 epochs with mini-batches of size $M=64$
and the following configuration. For the critic $d$, we use a 3-layer MLP with 10
hidden nodes per layer and ReLU activations. At each epoch, Adam is run for
$n_\text{critic}=5$ iterations with a step size $\alpha=0.01$ and otherwise default parameters.
Its inner first and second moment vectors are reset at
each outer epoch in order to avoid building momentum in staled directions.  For
estimating $\lambda^*$, we parametrize $\bftheta$ as $\log(\lambda)$ and use a univariate Gaussian proposal distribution
$q(\bftheta|\bfpsi)$ initialized with a mean at $\log(5)$ and unit variance. At
each epoch, parameters $\bfpsi$ are updated by taking one Adam step, with step size
$\alpha=0.01$. The gradient penalty coefficient is set to
$\lambda=0.025$, and the entropy penalty is evaluated at both $\gamma=0$ and $\gamma=1$.

The top left plot in Figure~\ref{fig:poisson} illustrates the resulting proposal
distributions $q(\bftheta|\bfpsi)$ after AVO.  For
both $\gamma=0$ and $\gamma=1$, the proposal distributions correctly concentrate
their density around the true parameter value $\log(\lambda^*) = 1.94$. Under
the effect of entropic regularization,
the proposal distribution for $\gamma=1$ concentrates its mass tightly,
yielding in this case precise inference.  The top right plot compares the
model distributions to the true distribution.  As theoretically expected from
adversarial training, we see that the resulting distributions align with
the true distribution, with in this case visually slightly better results for the penalized
model.  The bottom plot of Figure~\ref{fig:poisson} shows empirical estimates
of $-U_d$ with respect to the epoch number. For both $\gamma=0$ and $\gamma=1$,
the curves quickly fall towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\bftheta)}
[d(\tilde{\mathbf{x}};\bfphi)] \approx \mathbb{E}_{\mathbf{x} \sim
p_r(\mathbf{x})} [d(\mathbf{x};\bfphi)]$ and that the critic cannot distinguish
between true and model data. This confirms that adversarial variational optimization
works despite the discreteness of the data and lack of access to the
density $p(\bfx | \bftheta)$ or its gradient.

@@

As a second toy example, we consider a generator producing
5-dimensional continuous data, as originally specified in Section 4.2 of
\cite{cranmer2015approximating}.
% More specifically, we consider the following
% generative process:
% \begin{itemize}
% \item $\bfz = (z_0, z_1, z_2, z_3, z_4)$, such that \\
% $z_0 \sim {\cal N}(\mu=\alpha, \sigma=1)$,
% $z_1 \sim {\cal N}(\mu=\beta, \sigma=3)$,
% $z_2 \sim {\text{Mixture}}[\frac{1}{2}\,{\cal N}(\mu=-2, \sigma=1), \\  \frac{1}{2}\,{\cal N}(\mu=2, \sigma=0.5)]$,
% $z_3 \sim {\text{Exponential}(\lambda=3)}$, and
% $z_4 \sim {\text{Exponential}(\lambda=0.5)}$;
% \item $\mathbf{x} = R  \bfz$, where $R$ is a fixed
% semi-positive definite $5 \times 5$ matrix defining a fixed projection
% of $\bfz$ into the observed space.
% \end{itemize}
Again, AVO does not have access to the density or its gradient, only samples from the generative model.
We consider observed data generated at the nominal values $\bftheta^* = (\alpha^*=1,\beta^*=-1)$.
The simulator parameters are modeled with a factored  (mean field) Gaussian
proposal distribution $q(\bftheta|\bfpsi) = q(\alpha|\bfpsi) q(\beta|\bfpsi)$, where each component was
initialized with zero mean and unit variance.
Hyper-parameters are set to $M=64$, $n_\text{critic}=100$, $\lambda=0.025$, $\gamma=10$ and
Adam configured with a step size $\alpha=0.01$ and otherwise default parameters.
The architecture of the critic is the same as in the previous example.

Starting with a proposal distribution $q(\bftheta|\bfpsi)$ largely spread over
the parameter space, as illustrated in the left plot of Figure~\ref{fig:multi},
AVO quickly converges towards a proposal distribution whose density
concentrates around the nominal values $\bftheta^*$, as shown in the right plot of Figure~\ref{fig:multi}.
Overall, this example further illustrates and confirms the ability of adversarial
variational optimization for inference with multiple parameters and multidimensional
data, where reliable approximations of $p(\mathbf{x}|\bftheta)$ in a traditional
MLE setting would otherwise be difficult to construct.

@@

\subsection{Benchmarks}

\glnote{Compare against ABC and others.}
\glnote{Sections 6.1 and 6.4 of Bernton.}

@@

This work sits at the intersection of several lines of research related to
likelihood-free inference, approximate Bayesian computation (ABC),
implicit generative models, and variational inference.
Viewed from the literature around implicit generative models based on neural networks,
the proposed method can be considered as a direct adaptation of
generative adversarial networks~\cite{goodfellow2014generative,2017arXiv170104862A, 2017arXiv170107875A} to
non-differentiable simulators using variational optimization~\cite{2012arXiv1212.4507S,staines2013optimization}.

@@

Likewise, AVO closely relates to recent extensions of GANs, such as
Adversarial Learned Inference (ALI)~\cite{dumoulin2016adversarially},
Bidirectional GANs (BiGANs)~\cite{donahue2016adversarial},
$\alpha$-GAN~\cite{rosca2017variational}, Adversarial Variational Bayes
(AVB)~\cite{DBLP:journals/corr/MeschederNG17}, and the \texttt{PC-Adv}
algorithm of  \cite{2017arXiv170208235H}, which
add an inference network to the generative model.  Each of these assume a
tractable density $p(\bfx|\bftheta)$ that is differentiable with respect to
$\bftheta$, which is  not satisfied in the likelihood-free setting. Our lifting
of the non-differentiable simulator $p(\bfx|\bftheta)$ to the variational
program $q(\bfx | \bfpsi)$ provides the ability to differentiate expectations
with respect to $\bfpsi$ as in Eqn~\ref{eqn:approx-grad}; however, the density
$q(\bfx | \bfpsi)$ is still intractable. Moreover, we do not attempt to define a
recognition model  $q(\bfz, \bftheta|\bfpsi)$ as the latent space $\mathcal{Z}$
of many real-world simulators is complicated and not amenable to a neural
recognition model.

@@

Preliminary results on toy problems with discrete and continuous data validate
the proposed method. While the obtained results are encouraging, the complete
validation of the method remains to be carried out in real conditions on a full
fledged scientific simulator -- which we plan to achieve for a next version of
this work. In terms of method, several components need further investigation.
First, we need to better study the interplay between the entropic regularization and the
adversarial objectives. Second, we should better understand the dynamics of the
optimization procedure, in particular when combined with momentum-based
optimizers like Adam. Third, we need to consider whether less noisy estimates of
the gradients $\nabla_\bfpsi U_g$ can be computed.

@@

As a more realistic example, we here consider a (simplified) simulator from
particle physics for electron--positron collisions resulting in muon--antimuon
pairs ($e^+e^- \rightarrow \mu^+\mu^-$). The simulator approximates the
distribution of observed measurements $\mathbf{x} = \cos(A) \in [-1,1]$, where $A$ is the
polar angle of the outgoing muon with respect  to the originally incoming
electron. Neglecting measurement uncertainty induced from the particle detectors,
this random variable is approximately distributed as
\begin{equation}\label{eq:simple_weinberg}
p(\mathbf{x}|E_\text{beam}, G_f) = \frac{1}{Z} \left[ (1 + \mathbf{x}^2) + c(E_\text{beam}, G_f) \mathbf{x} \right]
\end{equation}
where $Z$ is a known normalization constant and $c$ is an asymmetry coefficient
function. Due to the linear term in the expression, the density $p(\mathbf{x} |
E_\text{beam}, G_f)$ exhibits a so-called {\it forward-backward} asymmetry.  Its
size depends on the values of the parameters $E_\text{beam}$ (the beam energy)
and $G_f$ (the Fermi constant) through the coefficient function $c$.
